Stochastic Variance-Reduced Policy Gradient

### 2018-02-08 svrpg.tex
minor fix in Lemma B6

%Lemma 0
\begin{restatable}[]{lemma}{auxzero}\label{lemma:aux0}
Under Assumption \ref{ass:bounded_score}, for any function $\varphi(\vtheta_t^{s+1})$ which is deterministic for a fixed $\vtheta_t^{s+1}$:
\begin{align*}
\left|\Ets[t]{\dotprod{\gradBlack{\vtheta_t^{s+1}}}{\varphi(\vtheta_t^{s+1})}}
-\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{\varphi(\vtheta_t^{s+1})}}
\right|
\leq
\frac{1}{2N}\Vars{g(\cdot\vert\wt{\vtheta}^s)} +\frac{1}{2}\Ets[t-1]{\norm[]{\varphi(\vtheta_t^{s+1})}^2}
\end{align*}
\end{restatable}
\begin{proof}
\begin{align}
	\Ets{\dotprod{\gradBlack{\vtheta_t^{s+1}}}{\varphi(\vtheta_t^{s+1})}}
	&=
	\Ets{\dotprod{\gradIdeal{\vtheta_t^{s+1}}}{\varphi(\vtheta_t^{s+1})}} +
	\Ets[t-1]{\dotprod{e_s}{\varphi(\vtheta_t^{s+1})}} \label{eq:6}\\
	&=
	\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{\varphi(\vtheta_t^{s+1})}} +
	\Ets[t-1]{\dotprod{e_s}{\varphi(\vtheta_t^{s+1})}} \label{eq:7}\\
	&=
	\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{\varphi(\vtheta_t^{s+1})}} \nonumber\\
	&\qquad+
	\dotprod{\Ets{e_s}}{\Ets{\varphi(\vtheta_t^{s+1})}}
	+\Covts[t-1]{\gradApp{\wt{\vtheta}^s}{N}}{\varphi(\vtheta_t^{s+1})}  \label{eq:new1}\\
	&= 
	\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{\varphi(\vtheta_t^{s+1})}} \nonumber\\
	&\qquad+
	\Covts[t-1]{\gradApp{\wt{\vtheta}^s}{N}}{\varphi(\vtheta_t^{s+1})} \label{eq:8}
\end{align}
where~\eqref{eq:6} is from Definition~\ref{def:ideal};~\eqref{eq:7} is from the fact that $\gradIdeal{\vtheta_t^{s+1}}$ is both unbiased and independent from $\varphi(\vtheta_t^{s+1})$ \wrt the sampling at time $t$ alone, which is not true for $\gradBlack{\vtheta_t^{s+1}}$;~\eqref{eq:new1} is from the fact that $\gradJ{\wt{\vtheta}^s}$ is constant \wrt $\Vars{\cdot}$;~\eqref{eq:8} is from $\Ets{e_s}=0$.
\todopir{In Eq~\ref{eq:7} the first expected value is on $t-1$? In any case even $t$ is fine.}
Hence:
\begin{align}
	\left|\Ets{\dotprod{\gradBlack{\vtheta_t^{s+1}}}{\varphi(\vtheta_t^{s+1})}}
	\right.&-\left.\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{\varphi(\vtheta_t^{s+1})}}\right| 
	=
	\left|\Covts[t-1]{\gradApp{\wt{\vtheta}^s}{N}}{\varphi(\vtheta_t^{s+1})}\right|  
	\nonumber\\
	&\leq
    \sqrt{\Vars{\gradApp{\wt{\vtheta}^s}{N}}} \cdot \sqrt{\Varts[t-1]{\varphi(\vtheta_t^{s+1})}} \label{eq:9a}\\
	%
	&\leq	
	\frac{1}{2}\Vars{\gradApp{\wt{\vtheta}^s}{N}} +\frac{1}{2}\Varts[t-1]{\varphi(\vtheta_t^{s+1})}\label{eq:9}\\
	%
	&=
	\frac{1}{2N}\Vars{g(\cdot\vert\wt{\vtheta}^s)} +\frac{1}{2}\Varts[t-1]{\varphi(\vtheta_t^{s+1})} \label{eq:10}\\
	%
	&\leq
	\frac{1}{2N}\Vars{g(\cdot\vert\wt{\vtheta}^s)} +\frac{1}{2}\Ets[t-1]{\norm[]{\varphi(\vtheta_t^{s+1})}^2},
	\nonumber
\end{align}
where~\eqref{eq:9a} comes from Cauchy-Schwarz inequality
\[
        |\Cov(X,Y)| = |\EVV[]{(X-\mu_X)\transpose{(Y-\mu_Y)}}| \leq \EVV[]{(X-\mu_X)^2}^{1/2}\EVV[]{(Y-\mu_Y)^2}^{1/2} = \sqrt{\Var(X) \Var(Y)},
\]
\eqref{eq:9} is from Young's inequality, and~\eqref{eq:10} is from the definition of $\gradApp{\vtheta}{N}$.
\todopir{comment the last step.}
\end{proof}

>>>>

%Lemma 0
\begin{restatable}[]{lemma}{auxzero}\label{lemma:aux0}
Under Assumption \ref{ass:bounded_score}, for any function $\varphi(\vtheta_t^{s+1})$ which is deterministic for a fixed $\vtheta_t^{s+1}$:
\begin{align*}
\left|\Ets[t]{\dotprod{\gradBlack{\vtheta_t^{s+1}}}{\varphi(\vtheta_t^{s+1})}}
-\Ets[t-1]{\dotprod{\gradJ{\vtheta_t^{s+1}}}{\varphi(\vtheta_t^{s+1})}}
\right|
\leq
\frac{1}{2N}\Vars{g(\cdot\vert\wt{\vtheta}^s)} +\frac{1}{2}\Ets[t-1]{\norm[]{\varphi(\vtheta_t^{s+1})}^2}
\end{align*}
\end{restatable}
\begin{proof}
\begin{align}
	\Ets{\dotprod{\gradBlack{\vtheta_t^{s+1}}}{\varphi(\vtheta_t^{s+1})}}
	&=
	\Ets{\dotprod{\gradIdeal{\vtheta_t^{s+1}}}{\varphi(\vtheta_t^{s+1})}} +
	\Ets[t-1]{\dotprod{e_s}{\varphi(\vtheta_t^{s+1})}} \label{eq:6}\\
	&=
	\Ets[t-1]{\dotprod{\gradJ{\vtheta_t^{s+1}}}{\varphi(\vtheta_t^{s+1})}} +
	\Ets[t-1]{\dotprod{e_s}{\varphi(\vtheta_t^{s+1})}} \label{eq:7}\\
	&=
	\Ets[t-1]{\dotprod{\gradJ{\vtheta_t^{s+1}}}{\varphi(\vtheta_t^{s+1})}} \nonumber\\
	&\qquad+
	\dotprod{\Ets[t-1]{e_s}}{\Ets[t-1]{\varphi(\vtheta_t^{s+1})}}
	+\Covts[t-1]{\gradApp{\wt{\vtheta}^s}{N}}{\varphi(\vtheta_t^{s+1})}  \label{eq:new1}\\
	&= 
	\Ets[t-1]{\dotprod{\gradJ{\vtheta_t^{s+1}}}{\varphi(\vtheta_t^{s+1})}} \nonumber\\
	&\qquad+
	\Covts[t-1]{\gradApp{\wt{\vtheta}^s}{N}}{\varphi(\vtheta_t^{s+1})} \label{eq:8}
\end{align}
where~\eqref{eq:6} is from Definition~\ref{def:ideal};~\eqref{eq:7} is from the fact that $\gradIdeal{\vtheta_t^{s+1}}$ is both unbiased and independent from $\varphi(\vtheta_t^{s+1})$ \wrt the sampling at time $t$ alone, which is not true for $\gradBlack{\vtheta_t^{s+1}}$;~\eqref{eq:new1} is from the fact that $\gradJ{\wt{\vtheta}^s}$ is constant \wrt $\Vars{\cdot}$;~\eqref{eq:8} is from $\Ets[\cdot]{e_s}=0$.
Hence:
\begin{align}
	\left|\Ets{\dotprod{\gradBlack{\vtheta_t^{s+1}}}{\varphi(\vtheta_t^{s+1})}}
	\right.&-\left.\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{\varphi(\vtheta_t^{s+1})}}\right| 
	=
	\left|\Covts[t-1]{\gradApp{\wt{\vtheta}^s}{N}}{\varphi(\vtheta_t^{s+1})}\right|  
	\nonumber\\
	&\leq
    \sqrt{\Vars{\gradApp{\wt{\vtheta}^s}{N}}} \cdot \sqrt{\Varts[t-1]{\varphi(\vtheta_t^{s+1})}} \label{eq:9a}\\
	%
	&\leq	
	\frac{1}{2}\Vars{\gradApp{\wt{\vtheta}^s}{N}} +\frac{1}{2}\Varts[t-1]{\varphi(\vtheta_t^{s+1})}\label{eq:9}\\
	%
	&=
	\frac{1}{2N}\Vars{g(\cdot\vert\wt{\vtheta}^s)} +\frac{1}{2}\Varts[t-1]{\varphi(\vtheta_t^{s+1})} \label{eq:10}\\
	%
	&\leq
	\frac{1}{2N}\Vars{g(\cdot\vert\wt{\vtheta}^s)} +\frac{1}{2}\Ets[t-1]{\norm[]{\varphi(\vtheta_t^{s+1})}^2},
	\nonumber
\end{align}
where~\eqref{eq:9a} comes from Cauchy-Schwarz inequality
\[
        |\Cov(X,Y)| = |\EVV[]{(X-\mu_X)\transpose{(Y-\mu_Y)}}| \leq \EVV[]{(X-\mu_X)^2}^{1/2}\EVV[]{(Y-\mu_Y)^2}^{1/2} = \sqrt{\Var(X) \Var(Y)},
\]
\eqref{eq:9} is from Young's inequality, and~\eqref{eq:10} is from the definition of $\gradApp{\vtheta}{N}$.
\todopir{comment the last step.}
\end{proof}
